\documentclass[a4paper,11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{array}
\usepackage{ragged2e}
\usepackage{caption}

\title{Bill of Materials — Minor Smart Things}
\date{\today}

\begin{document}
\maketitle

\bigskip
\small
\begin{longtable}{@{}p{0.07\linewidth} p{0.18\linewidth} p{0.18\linewidth} p{0.37\linewidth} p{0.20\linewidth}@{}}
\caption{Bill of Materials}\\
\toprule
Qty & Item & Model / Part number & Description & Notes \\
\midrule
\endfirsthead

\toprule
Qty & Item & Model / Part number & Description & Notes \\
\midrule
\endhead

\midrule
\multicolumn{5}{r}{\footnotesize Continued on next page} \\
\endfoot

\bottomrule
\endlastfoot

1 & 3144 Hall Effect Sensor & A3144 & Unipolar Hall-effect switch for detecting magnetic fields/rotation; through-hole TO-92 package (common). & TO-92 package (common) \\
1 & 0.96 inch OLED Display 128*64 pixels Blauw - I2C & SSD1306 & 0.96" 128x64 monochrome OLED module, I2C interface (blue). & I2C address may vary (0x3C/0x3D) \\
1 & MG90 9g Metal-geared Hobby Servo & MG90 & Metal-geared 9g class hobby servo, 5\,V supply; improved torque and durability for sunshade actuation. & Metal gears; ensure 5\,V supply capable of servo peaks \\
1 & BH1750 16bit Digitale I2C Licht Sensor Module & BH1750 & Digital ambient light sensor, I2C, 16-bit resolution. & Typical breakout module \\
1 & DHT22 & DHT22 & Digital temperature and humidity sensor (single-wire protocol). & Check operating voltage and wiring \\
1 & PCB & Custom-made & PCB for mounting ESP32, sensors and connectors; includes headers for sensors, servo connector, and mounting holes. & Gerber files provided; double-sided, 2.54mm headers; silk labeling \\
20 & jumper cables & - & Jumper cable set (male-female) for prototyping. & Various lengths, mostly 10cm; quality varies \\
8 & 3D products & - & 3D-printed parts (PLA/PETG) such as brackets and mounts. & Specify material, orientation, tolerances \\
2 & Esp32 & ESP32 (e.g., WROOM-32) & Wi-Fi + Bluetooth microcontroller module. & Verify module form factor and flash size \\
3 & External source & Assorted passive components & Resistors, capacitors, header pins, connectors, screws and standoffs required for assembly. & Refer to schematic for values and quantities \\
1 & Magnet & - & Neodymium magnet — size and grade TBD. & Choose size/strength according to sensor/mounting \\
1 & Ø6 mm tube & - & Aluminium tube, length: 10 cm & Material: Aluminium \\

\end{longtable}
\normalsize

\end{document}